---
title: "practice analysis"
author: "ailani"
date: "9/7/26"
output: html_document
---

Load packages
```{r, warning = FALSE, message = FALSE}

#list of packages (Diveica et al., 2025)
packages <- c("tidyverse", "bootnet", "NetworkComparisonTest", "qgraph", "igraph", "papaja", "ggcorrplot", "rstatix", "psych", "ggrepel", "ggExtra")

#load each package
for (pkg in packages) {
  library(pkg, character.only = TRUE, quietly=TRUE)
}

# notin operator (Diveica et al, 2025)
'%notin%' <- negate('%in%')
```

R environment, version info for reproducibility 
```{r session-info, include=TRUE}
sessionInfo() 
```

Load datasets 
``` {r datasets}
setwd('/Users/ailani/Projects/iconic-network-analysis/data/')

AOA <- read_csv('kuperman_2012_AOA.csv')
freq <- read_csv('brysbaert_2012_SUBTLEX_POS.csv')
lanc <- read_csv('Lancaster_sensorimotor_norms_for_39707_words.csv')
spec <- read_csv('muraki_2026_specificity.csv')
emot <- read_csv('warriner_2013_affective_norms.csv')
icon <- read_csv("winter_2024_iconicity_ratings_cleaned.csv")

```


Select relevant columns from each dataset
```{r}
AOA <- AOA %>% select(Word, Rating.Mean) %>% 
  rename(AOA = Rating.Mean)

freq <- freq %>% 
  rename(Frequency = FREQcount) %>% 
  select(Word, Frequency)

lanc <- lanc %>% select(Word, Auditory.mean:Torso.mean) %>% 
  mutate(Word = tolower(Word)) %>%
  rename_with(~ str_remove(.x, "\\.mean"), Auditory.mean:Torso.mean)

spec <- spec %>% select(Word, Value) %>%  
  rename(Specificity = Value) #make sure if Value is the right column to use

emot <- emot %>% select(Word, A.Mean.Sum, V.Mean.Sum) %>% 
  mutate(Valence = abs(V.Mean.Sum - 5)) %>% # Get the valence extremity value
  rename(Arousal = A.Mean.Sum) %>% 
  select(Word, Arousal, Valence)


icon <- icon %>% select(word, rating, rating_sd)  %>% #make sure if rating is right column 
  rename(Word = word, Iconicity = rating, Iconicity_SD = rating_sd) 
```


Join to full dataset 
```{r}
df <- reduce(list(AOA, lanc, spec, emot, icon, freq), inner_join, by = "Word")
```

See how many words overlapped
```{r}
nrow(df)
```

There are `r nrow(df)` overlapping words across all the data sets representing the properties of interest.  

Log-transform frequency data
```{r}
# First make sure NAs in Frequency are 0 
df <- mutate(df, 
  Frequency = ifelse(is.na(Frequency), 0, Frequency))

df <- mutate(df, 
      LogFreq = log10(Frequency + 1))
```

Standardize variables (Z-scores)
```{r}
z_score <- function(x) (x - mean(x, na.rm = TRUE)) / sd(x)
df <- df %>% mutate(across(c(AOA, Auditory:Torso, Specificity, Arousal, Valence, LogFreq), z_score, .names = "{.col}_z"))
```


Split into two data frames: one with high iconicity words and another with low iconicity (more arbitrary) words. For lack of conventions in iconicity literature, use tertile split.
```{r}
# Find the tertile cutoffs on iconicity ratings 
tertiles <- quantile(df$Iconicity, probs= c(1/3, 2/3), na.rm=TRUE)
cut_low <- tertiles[1]
cut_high <- tertiles[2]

# Split into high-iconicity and low-iconicity word groups 
high_iconic <- df %>% filter(Iconicity >= cut_high)
low_iconic <- df %>% filter(Iconicity <= cut_low)
```

Check group size 
```{r}
nrow(high_iconic)
nrow(low_iconic)
```

There are more high_iconic words (`{r} nrow(high_iconic)`) than low_iconic (`{r} nrow(low_iconic)`), so match to low_iconic 
```{r}
n_match <- min(nrow(high_iconic), nrow(low_iconic))

high_iconic <- high_iconic %>% 
  arrange(desc(Iconicity)) %>% 
  slice_head(n=n_match) %>%
  mutate(Iconic = "Iconic")

low_iconic <- low_iconic %>% 
  arrange(Iconicity) %>% 
  slice_head(n=n_match) %>%
  mutate(Iconic = "Non-iconic")
```

Check if the rows of the split data sets are equal, along with their respective means and ranges 
```{r}
# Check  
nrow(high_iconic) == nrow(low_iconic)
mean(high_iconic$Iconicity); range(high_iconic$Iconicity); sd(high_iconic$Iconicity)

mean(low_iconic$Iconicity);  range(low_iconic$Iconicity); 
sd(low_iconic$Iconicity)
```

Also get the mean of the SDs (inter-rater variability) for each group
``` {r}
high_iconic %>% summarise(mean_sd = mean(Iconicity_SD, na.rm=TRUE))
low_iconic %>% summarise(mean_sd = mean(Iconicity_SD, na.rm=TRUE))

df %>% summarise(mean_SD_full = mean(Iconicity_SD, na.rm=TRUE))
```

Now, both high iconic group and low iconic group have same number of words, `{r} nrows(high_iconic)`.

The mean iconicity of the high_iconic set is `{r} mean(high_iconic$Iconicity)` and the range is `{r} range(high_iconic$Iconicity)` 

The mean iconicity of the low_iconic set is `{r} mean(low_iconic$Iconicity)` and the range is `{r} range(low_iconic$Iconicity)` .

Mean inter-rater SD was 1.78 (High) and 1.74 (Low). These are both slightly lower than full-sample mean SD of 1.81. 
```{r}
t.test(df$Iconicity_SD, mu = 1.775)  # 1.775 = combined High/Low mean, adjust as needed

middle_icon_words <- df %>% 
  filter(Iconicity > cut_low, Iconicity < cut_high)

mean(middle_icon_words$Iconicity_SD, na.rm = TRUE)
```


Combine the split data sets into one (removing the middle iconicity words)
 ```{r}
data_both <- bind_rows(high_iconic, low_iconic) %>%
  select(Word, Iconic, everything()) %>%
  mutate(Iconic = factor(Iconic, levels = c("Iconic", "Non-iconic")))
```

Univariate group comparison of semantic dimensions between high iconic vs low iconic words (`broom` package)
```{r}
# t-tests to compare ratings between high-iconic vs low-iconic 
# list of variables
vars_test <- c("AOA", "Auditory", "Gustatory","Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", "Head", "Mouth", "Torso", "Specificity", "Arousal", "Valence")

results <- map_dfr(vars_test, function(var) {
  formula <- as.formula(paste(var,"~ Iconic"))
  t_result <- t.test(formula, data=data_both)

  tidy(t_result) %>%
    mutate(Variable = var) %>% 
    select(Variable, estimate, statistic, p.value, conf.low, conf.high)
})
```

Adjust for multiple comparisons with Bonferroni correction (following Diveica et al., 2025)
```{r}
results <- results %>% 
  mutate(p_adj = p.adjust(p.value, method="bonferroni")) %>%
  arrange(p_adj)

print(results)
```

Calculate the effect sizes (Cohen's d) 
```{r}
effect_sizes <- map_dfr(vars_test, function(var){
  formula <- as.formula(paste(var, "~ Iconic"))
  d <- cohens_d(formula, data=data_both)
  tibble(Variable = var, effsize= d$effsize, magnitude = d$magnitude)
})

results <- results %>%
  left_join(effect_sizes, by = "Variable")

print(results)
```

Visualize the differences with violin plots
```{r}
# Reshape data to long format
both_long <- data_both %>% 
  select(Iconic, all_of(vars_test)) %>%
  pivot_longer(-Iconic, names_to = "Variable", values_to = "Value")

# Build plot 
ggplot(both_long, 
    aes(x = Iconic, y = Value, fill= Iconic)) +
    geom_violin(alpha=0.5) +
    geom_boxplot(width=0.15, outlier.shape = NA) +
    facet_wrap(~ Variable, scales= "free_y") +
    theme_minimal()


```


Correlation martix for full data set 
``` {r}
library(dplyr)
cor_data <- df %>%
  select(all_of(vars_test), Iconicity)

cor_matrix <- cor(cor_data, use="pairwise.complete.obs", method = "pearson")
print(round(cor_matrix,2))

# get p-values for correlations 
library(Hmisc)

rcorr_result <- rcorr(as.matrix(cor_data), type = "pearson")

rcorr_result$r  # correlation coefficients
rcorr_result$P  # p-values

# Visualize
library(corrplot)
corrplot(cor_matrix, 
         method = "color",
         type = "upper",
         #order = "hclust",
         p.mat = rcorr_result$P,
         sig.level = 0.05,
         insig = "blank",       # blank out non-significant cells
         addCoef.col = "black",
         number.cex = 0.6,
         tl.col = "black",
         title = "Full",
         tl.srt = 45,
         diag = FALSE)
```

Separate correlation matrices for high vs low iconic groups 
``` {r} 
cor_hi_data <- high_iconic %>%
  select(all_of(vars_test), Iconicity)
cor_low_data <- low_iconic %>%
  select(all_of(vars_test), Iconicity)

cor_high <- cor(cor_hi_data, use = "pairwise.complete.obs")
cor_low  <- cor(cor_low_data,  use = "pairwise.complete.obs")

corrplot(cor_high, method = "color", type = "upper", addCoef.col = "black", 
         number.cex = 0.5, tl.col = "black", tl.srt = 45, diag = FALSE,
         title = "High iconicity", mar = c(0,0,2,0))

corrplot(cor_low, method = "color", type = "upper", addCoef.col = "black", 
         number.cex = 0.5, tl.col = "black", tl.srt = 45, diag = FALSE,
         title = "Low iconicity", mar = c(0,0,2,0))
```